\large{\bf Large effects and the infinitesimal model }

### 2023-12-03 review-responses.tex

\vskip 2em

\noindent
{\bf To the Editor(s) -- }

\vskip 1em

We are writing to submit a revised version of our manuscript,
``Large effects and the infinitesimal model''.
In response to the reviewer's comments
we have made quite a few additions to the text
which we hope will make the paper more accessible to a wider audience.
We hope you agree!

Along with our revised paper and the responses to reviewers,
we are also submitting a PDF that has both changes to the paper since the first submission highlighted,
and hyperlinked reviewer comments.
(Line numbers in the standalone responses to reviewers
refer to the PDF of the paper without changes highlighted,
while line numbers in the merged document with changes highlighted are self-consistent.)

@@

\pagebreak

%%%%%%%%%%%%%%
\section*{AE comments:}

\begin{quote}
The two reviewers find the results interesting and worth being eventually published in the special issue. However, they highlight a number of places in the text where the properties claimed or the precise questions addressed need serious clarification. As the readership of the special issue will be quite broad, it is important that the main messages are accessible to colleagues who do not necessarily know Fisher's infinitesimal model precisely. Therefore, although the mathematical results in themselves do not require deep modifications, I recommend that a major revision of the paper should be provided, which would take into account the different comments and suggestions of the reviewers.
\end{quote}

Thanks very much to the reviewers for their comments.
We agree that our initial version of the manuscript didn't have as broad an envisioned audience
as it should have.
We've tried to improve this, by adding additional context and discussion --
in particular, in Sections 3 and 4, and to the Discussion.

@@

Thanks for the careful reading and useful comments.
We hope you like our revisions.

\begin{point}{}
    Paragraph starting ``This suggests exploring whether stable distributions..'' \revref:
The questions as stated: e.g., whether stable distributions can ``stand in'' for a Gaussian distribution in the infinitesimal model or whether the results of Barton et al ``carry over'' when the sum of effect sizes follows a stable law are quite vague and allow for multiple interpretations. Ideally, in order to address whether Barton et al's results (do not) carry over, one would like to at least show that with a power law distribution of effect sizes, the ``Mendelian noise term'', i.e., the difference between the offspring trait value and the midparent value, is (not) independent of the midparent value in the limit of a very large number of loci. And indeed this is what the authors address
somewhat non-rigorously (and for the case of a Cauchy distribution of effect sizes) in section 5 and using simulations : see e.g., figures 4C and 4f, figures 6C and 6F. Unsurprisingly, these simulations show that with e.g., Cauchy distribution of effect sizes, the non-independence between the segregation/noise term and the midparent trait value does not hold. However, a lot of the mathematical analysis in the paper (sections 3 and 4) explores when the distribution of trait values in a population is (not) Gaussian, which is a completely different question.
\end{point}

\reply{
    We agree with most of this statement --
    the best answer to the admittedly vague question we pose here
    does have to do with whether the Mendelian noise term can be taken to be independent of parental values,
    and we do explore some other -- we think related and interesting -- questions along the way.
    However, we think there is some confusion here about both Sections 3 and \ref{sec:abnormal},
    which we discuss more below: in particular, the focus of Section 3
    is not on the population distribution,
    but instead on the very question brought up here
    (whether the Mendelian term can be independent of the parental values).
    However, the assumptions and motivations were not as clear as they could have been
    in the previous draft, and this is hopefully more clear now --
    see responses to points below.
}

@@

\begin{point}{\revref}
In the same paragraph: I don't think the claim that ``independence of offsprings' deviations and midparent values implies a Gaussian distribution'' (which is also stated in the abstract) is true? Surely, ``independence of offsprings' deviations and midparent values'' is just the basic statement of the infinitesimal model (say, in Barton et al 2017) and the infinitesimal model does not imply a Gaussian distribution of trait values? See also section 3.
\end{point}

\reply{
    We have modified the sentence to say
    ``implies a Gaussian distribution for the offsprings' deviations'',
    to make it clear that this statement does not refer to the population distribution,
    but rather to the Mendelian sampling term \revref.
    See additional points below on this topic.
}

@@

\begin{point}{}
I was also wondering about the ``goodness of fit'' of observed distributions to power laws. I am not necessarily recommending a detailed quantitative analysis, but nevertheless it may be useful to comment on whether one has power to say reject alternative distributions (e.g., a mixture of exponentials) that are often considered. If not, then perhaps the exact exponents inferred are unimportant, and what the data points to more broadly, is a broad distribution of effect sizes.
\end{point}

\reply{
    This is a good question, and we had in fact done some exploration of this question
    but not included it in the paper.
    We've still not done a detailed analysis,
    but now provide some discussion and numerical exploration of our precision here:
    see Figure~\ref{fig:power} and \revref.
}

\begin{point}{}
    Section 3, Sentence starting ``We know from \citet{barton2017infinitesimal}, \ldots'' \revref: As far as I understand, the statement of \citet{barton2017infinitesimal} is that the joint distribution of trait values of offspring is multivaraite normal, conditioned on the two parents. I don't think the joint distribution of two randomly picked parents and their offspring is mutivariate normal, as this would require some kind of normality assumption about trait values in the population as a whole. It'd be helpful to distinguish more carefully between conditioned and unconditioned distributions, especially in this section.
\end{point}

\reply{
    Looking back at \citet{barton2017infinitesimal}, we see where the confusion comes from,
    and have added some clarifying remarks at the end of this paragraph \revref.
    % Quoting \citet{barton2017infinitesimal},
    % ``Even though the trait distribution across the whole population may be far from Gaussian,
    % the multivariate normal will play a central r\hat{o}le in our analysis''.
    In more detail:
    \citet{barton2017infinitesimal} (``main result'', equation 9)
    shows that conditioned on the trait values in generation $t$,
    the values of generation $t+1$ are also multivariate Gaussian.
    This implies that the joint distribution of traits of any subset of individuals
    is multivariate Gaussian (since the initial generation is Gaussian)
    -- as \citet{barton2017infinitesimal} says,
    \textit{``equally we could have conditioned on the values of any subset of relatives
    and the same would hold true: expected trait values would be
    a linear functional (determined by the pedigree) of the values
    on which we conditioned, but segregation variances would be unchanged.''}
    (More generally, if $X$ is multivariate Gaussian, and $Y$ conditioned on $X$ is also multivariate Gaussian
    with mean but not covariance depending on $X$,
    then $(X,Y)$ are jointly multivariate Gaussian.)
    We now also discuss the point further in a subsequent paragraph \llname{pop_not_gaussian}.
}

@@

\begin{point}{}
    In general, I find this section rather confused: the infinitesimal model basically implies that the ``Mendelian segregation term'' $R_{M}$ (which is the difference between an individual and the midparent value $\bar{Z}_{M}$) is normally distributed with mean zero and variance that depends on the relatedness between parents but is independent of the midparent value $\bar{Z}_{M}$. However, in Proposition \ref{prop:parental_contribs}, the authors claim that $\bar{Z}_{M}$ and $R_{M}$ are jointly Gaussian (which is equivalent to saying that the trait value of an individual and the mean of its parents' trait values is jointly Gaussian), which is in general not true.
\end{point}

\reply{
    If the reviewer had said ``confusing'' rather than ``confused'',
    then we would heartily agree.
    After some difficult thinking through of things,
    we've provided a good bit of additional clarification
    about what the Proposition means and doesn't mean:
    see in particular \llname{pop_not_gaussian}.
}

@@

\begin{point}{}
    I think it is also not correct to claim that ``Since models can be easily set up for which trait distributions are not Gaussian, the implication of this is that for such models, independence of the Mendelian sampling term is not likely a good assumption.'' \revref. As far as I understand, a non-Gaussian distribution of trait values can arise quite easily under the infinitesimal model (where the Mendelian segregation term is independent of the midparent value) e.g, due to strong or non-Gaussian selection or migration or some combination of the two (see, e.g., Figure 1 in Barton et al).
\end{point}

\reply{
    Thanks for the catch -- this was an unfortunate mistake that might have led to some of the previous confusion.
    We've changed this to ``since models can be easily set up for which the distribution
    of Mendelian sampling terms is not Gaussian, \ldots''
    and provided additional context \revref.
}

@@

\begin{point}{}
    Introduction: Also worth citing Fisher, Bulmer etc. (the original references) when introducing the infinitesimal model?
\end{point}

\reply{
    Good idea, we have added some citations \revref.
    However, note that what exactly are the ``original references'' is not so clear:
    ``\textit{Indeed, its origins are lost in the mists of time (M. Bulmer, W.G. Hill, pers.  comm.).}''
    \citep[from][]{barton2017infinitesimal}
}

\begin{point}{\revref}
    ``Perhaps the distribution of effect sizes within each gene is Normal'': I am confused by this. Do you mean that the distribution of effect sizes across all genes (with roughly the same ``proximity'' to the trait) is normal. If yes, then the above phrasing is a bit misleading. If what is meant that is really that the effect sizes of different alleles within a gene are normally distributed, then I am not sure if this is entirely plausible: see various papers by Turelli (e.g., ``Heritable Genetic Variation via Mutation-Selection Balance: Lerch's Zeta Meets the Abdominal Bristle.''). Presumably, a normal distribution of effect sizes for a sufficiently large genomic region is a good approximation but it is unclear what is large enough: smaller or larger than a typical gene...?
\end{point}

\reply{
    We've rephrased this, and replaced ``gene'' with ``region'' \revref --
    really, we just want to motivate a mixture model,
    as the "scale mixture of Gaussians" is a common way to get fat-tailed distributions,
    and at least roughly biologically plausible.
    On the other hand, we don't want to get into the weeds with more detail or discussion here.
}

@@

\begin{point}{}
    Figure \ref{fig:exponent_hist}: Is there also a correlation between the number of cases and the number of SNPs? In other words, to what extent are (b) and (c) independent?
\end{point}

\reply{
    Yes; we've added this to Figure~\ref{fig:exponent_hist}.
}

\begin{point}{\revref}
    Sentence starting ``To see what knowing the parental trait value...'': In what sense is this the ``parental'' trait value??
\end{point}

\reply{
    They are a member of the (first) ancestral generation, and therefore a (potential) parent.
    We've clarified this \revref.
}

@@

\begin{point}{}
    Discussion: In the context of theory combining large and small-effect loci, \citet{chevin2008selective}: ``Selective Sweep at a Quantitative Trait Locus in the Presence of Background Genetic Variation'' and \citet{lande1983response}: ``The response to selection on major and minor mutations affecting a metrical trait'' may also be somewhat relevant.
\end{point}

\reply{
    Thanks! We've incorporated these \revref.
}

%%%%%%%%%%%%%%
\reviewersection{2}

\begin{quote}
The authors investigate the possibility of extending results of \citet{barton2017infinitesimal} on the
justification of the infinitesimal model to distributions of effect sizes with a large tail (i.e.,
polynomial decay with exponent $\alpha < 2$). This is biologically interesting and important
because there are clear indications that variation in some traits is caused by a mixture
of polygenic (infinitesimal) contributions and one or few loci of large effect. It also a
mathematically difficult topic because it leads into the realm of generalized Central Limit
Theorems, namely stable laws.

@@

We're glad the reviewer liked this section,
but perhaps our introduction to the section was misleading --
in fact we do \emph{not} assume a Gaussian population distribution,
or any population distribution.
Instead, we \emph{ask} which distributions might be population distributions
under this generalized model,
using arguments similar to those found in the literature.
We've tried to make this more clear -- see additional text at \revreffull{1}{8}
and \llname{clarification3}
(and our response to Reviewer 1, point 8).

\begin{quote}
The authors use a simplistic Moran model to investigate these questions. However, mutation 
doesn’t seem to be modeled explicitly. Is it mutation if a parent is replaced by an
offspring of the type described in the display equation at \llname{offspring_eqn}?
Maybe yes, but not necessarily because of segregation and recombination contribute substantially to $R$.
Doesn’t the present assumption imply that the mutation rate is one and that mutation is modeled as a
kind of diffusion? In the latter case, a Gaussian distribution can result. With explicit
mutation models, where mutations occur per locus and a finite number of loci determines the
trait additively, a Gaussian distribution never results (and in no known limit; see above).
How does population size enter? Is this a deterministic Moran model?
\end{quote}

Mutation would enter into $R$, a point we now make \llname{mutation_note},
but without detail, as we are aiming here for a high-level and quick calculation,
not a detailed investigation.
In fact, we thought a bit about how to set up an explicit genotype-to-phenotype map
in which the Mendelian sampling terms are non-Gaussian and independent of parental traits,
but did not arrive at a concrete solution --
we think this is probably possible,
but as we know from Proposition~\ref{prop:parental_contribs},
cannot be a simple additive model.
Investigating how exactly a concrete model of mutation would affect things is beyond the scope of this paper,
and we hope the interesting bits of this section
outweigh the lack of an explicit mutationi model.
We have added a note that hopefully clarifies the questions
about population size \llname{deterministic_point}.

### 2023-12-03 writeup.tex
edit through

\begin{abstract}
    The infinitesimal model of quantitative genetics
    relies on the Central Limit Theorem to stipulate that
    under additive models of quantitative traits
    determined by many loci having similar effect size,
    the difference between an offspring's genetic trait component
    and the average of their two parents' genetic trait components
    is Normally distributed and independent of the parents' values.
    Here, we investigate how the assumption of similar effect sizes
    affects the model:
    in particular, if the tail of the effect size distribution is polynomial
    with exponent $\alpha < 2$, then sums of effects
    should be well-approximated by a ``stable distribution'',
    and find tail exponents between 1 and 2 in 
    effect sizes estimated by genome-wide association studies
    of many human disease-related traits.
    We show that the independence of offspring trait deviations from parental averages
    in many cases implies a Gaussian distribution,
    suggesting that non-Gaussian models of trait evolution must explicitly
    track the underlying genetics, at least for loci of large effect.
    We also characterize possible limiting trait distributions
    of the infinitesimal model with infinitely divisible noise distributions,
    and compare our results to simulations.
\end{abstract}

@@

The Gaussian distribution appears thanks to ``the'' Central Limit Theorem,
i.e., the universality of adding up a great many very small and independent things.
However, there is a wider class of Central Limit Theorems,
which describe the distributions of sums of a great many independent and exchangeable things:
the \emph{stable laws} \citep{gnedenko1968limit}.
Roughly speaking, these say that if we add together $n$ independent copies of a random variable $X$
for which $\P\{ X > t \}$ is proportional to $t^{-\alpha}$ for some $0 < \alpha < 2$,
then -- almost regardless of the details -- the result, scaled by $n^{1/\alpha}$,
is well-approximated by a universal form, an $\alpha$-stable distribution
(and the approximation becomes exact, as $n \to \infty$).
Oth the other hand, if $\alpha \ge 2$ then the distribution has finite variance,
and so the sum of $n$, divided by $\sqrt{n}$, converges to a Gaussian distribution.
In such ensembles, single large entries can be important:
concretely, the largest of the $n$ copies will be of order $n^{1/\alpha}$,
so that e.g., for $\alpha=1$ a positive proportion of the sum of $n$ entries will be contributed by a single one.
On the other hand, the largest of $n$ Gaussians is of order $\sqrt{\log n}$. \revpoint{2}{1} \revpoint{1}{11}

This suggests exploring whether stable distributions might reasonably stand in for the Gaussian
in the infinitesimal model. \revpoint{1}{1}
In this paper, we make preliminary explorations in that direction:
Do the results of \citet{barton2017infinitesimal}
carry over when the distribution of effect sizes falls in the domain of attraction of a stable law?
The immediate answer to this will be ``no'' --
in fact, we will see that independence of offsprings' deviations and midparent values \emph{implies} a Gaussian distribution for the offsprings' deviations, in some sense. \revpoint{1}{2}
This independence is a key reason why the infinitesimal model is tractable,
and so a conclusion is that additive models with non-Gaussian effects may not be well-approximated
by ``trait-only'' models that do not explicitly represent underlying genetic variation.
Having thus shown that ``infinitesimal models with non-Gaussian noise'' is on weak footing as a model of reality,
we then go on to analyze exactly these models.
(As motivation,
we have not shown that such models are \emph{not} good approximations to reality,
and it is interesting and perhaps informative to understand how the noise distribution affects results.)

@@

The ways in which genetic variants affect an organisms' trait can be complex,
and interactions between alleles, between loci, and with the environment are often important in practice.
In fact, many traits that have been claimed to be mostly genetically determined \revpoint{1}{14}
only appear heritable through non-genetic means,
e.g., systematic inequalities in resource distribution partitioned by class or race
\citep[e.g.,][]{davenport1916hereditary}.
In this paper, we follow the literature in working with the relatively simple purely additive case
of a trait that can be decomposed into a sum across loci of the contributions of the alleles from each locus,
even omitting an independent effect of ``the environment''.
However, we emphasize that such models should not always be the default for understanding biological variation,
especially for human traits.

@@

Since the Mendelian sampling term is a sum of terms of the form
$$
    (X_\ell - 1/2) \eta_\ell^{[1]} + (1/2 - X_\ell) \eta_\ell^{[2]},
$$
where $\eta_\ell^{[i]}$ is the effect of the allele in parent $i$ at locus $\ell$,
the appropriate central limit theorem for the sum is determined by
the tail behavior of $|(X - 1/2) \eta|$,
where $X$ is Bernoulli(1/2) and $\eta$ is the effect
of a uniformly chosen one of the two alleles from a uniformly chosen parent
at a uniformly chosen locus.
Since $|X-1/2| = 1/2$, we'd like to know
if $\P\{|X|>2t\}$ is proportional to $t^{-\alpha}$,
i.e., the value of
$$
    \lim_{t \to \infty} \frac{ - \log \P\{ |X| > 2 t \} }{ \log t } = \alpha ,
$$
if the limit exists.
Let $p_\ell$ be the minor allele frequency at the $\ell^\text{th}$ locus;
since the data we have assigns effect 0 to the major allele,
$\eta_\ell = 0$ with probability $1-p_\ell$.
Therefore, if we have estimated effects for $M$ SNPs
with minor allele frequency $p_\ell$ and effect size $e_\ell$ for the minor allele,
for $1 \le \ell \le M$,
then
$$
    \P\{ |\eta| > 2t \} \approx \frac{1}{M} \sum_{\ell : |e_\ell| > 2t} p_\ell .
$$
Note that the relevant effect size distribution here is weighted by frequency, \revpoint{1}{3}
which is necessary because we're looking at the deviation of a typical offspring from their parents,
and the chance a given allele is carried by a random parent is given by the frequency.
This will make a difference as frequencies and effect sizes are often correlated in practice.
So, to obtain a per-trait estimate of $\alpha$,
we plot $\log \frac{1}{M} \sum_{\ell : |e_\ell| > t} p_\ell$ against $\log t$
for a sequence of values of $t$,
and fit a linear model to the tail.
(To do this,
we take as the ``tail''
all values of $t$ for which the number of SNPs with absolute effect above $t$
is at least 20 and at most 10\% of all SNPs.) \revpoint{2}{3}
Examples are shown in Figure~\ref{fig:example_snps}.

@@

A possible concern is that tail exponents are difficult to estimate accurately, \revpoint{1}{4}
but simulation-based estimates of accuracy (Supplemental Figure~\ref{fig:power})
suggest that our estimates are unbiased but only accurate to within roughly $\pm 0.4$,
so the distribution of Figure~\ref{fig:exponent_hist} should not be taken as
a direct estimate of the underlying distribution of $\alpha$ values,
but is robust evidence for exponents strictly between 1 and 2 (and probably around 1.5).

@@

The Kac-Bernstein theorem states that if two independent random variables
have the property that their sum and difference are independent of each other,
then the random variables have a joint Gaussian distribution
(\citet{kac1939characterization,bernstein1941property},
and see \citet{kagan1973characterization} for many much more general results).
Since $R_M$ is (half) the difference and $\bar Z_M$ is (half) the sum
of $Z_M$ and $V_M$, and $Z_M$ and $V_M$ are independent,
if we assume that also $R_M$ and $\bar Z_M$ are independent,
then both are therefore jointly Gaussian.

@@

We should now discuss what this counter-intuitive result says and does not say.
The proposition says, roughly, that if the parental alleles are unlinked and independently chosen,
and meiotic segregation is fair,
then independence of Mendelian sampling term from midparent value
\emph{implies} a joint Gaussian distribution for the two.
We have not assumed anything about the distribution of the trait in the population:
rather, we have showed that any model of this sort
that assumes independence of Mendelian sampling term and midparent value
must also, by logical necessity, have a Gaussian distribution for the two.
In other words,
since models can be easily set up for which the distribution of Mendelian sampling terms
is not Gaussian,
the implication of the proposition
is that for such models, independence of the Mendelian sampling term
is not likely a good assumption. \revpoint{1}{7}
Note that the proof does not depend on independence of the $X_\ell$:
indeed, alleles may be inherited together,
although in practice this would quickly create dependencies between the effects
of alleles at such linked loci. \revpoint{2}{5}
Also note that it is \emph{not} usually true that the components --
for instance, $U_M^{[1]}$ and $U_M^{[2]}$ --
are independent of each other.
Finally, note that although they may not be independent,
the Mendelian sampling term $R_M$ and the midparent value $\bar Z_M$
\emph{are} always uncorrelated, since $Z_M$ and $V_M$ are exchangeable.

@@

Finally, a possible point of confusion is that \revpoint{1}{6}
the infinitesimal model is often phrased as a statement about offspring trait distributions
\emph{given} parental trait values,
while our statement above regards parental trait values as random.
However, the notions are equivalent:
the statement that
``for each pair of parents the deviation of the offspring's trait from midparent
has a distribution that does not depend on the parents' values''
is essentially the definition of statistical independence
of offspring's Mendelian sampling term and parental trait values.

@@

As observed in \cite{turelli2017commentary},
there are at least three interpretations of what is called ``the infinitesimal model''. 
The maximalist of these, what \cite{turelli2017commentary} calls the Gaussian-population approximation,
assumes parental trait-independent Gaussian Mendelian sampling terms,
and, moreover, that  population trait values are in deterministic mutation-selection balance, with a Gaussian empirical distribution
(\ie the \emph{fraction} of the population with trait value in $[z,z+dz)$
is approximately $\frac{1}{2\pi\sigma^{2}} e^{-(z-m)^{2}/(2\sigma^{2})} dz$
for suitable choices of the empirical mean and variance $m$ and $\sigma^{2}$).
As also discussed in \citet{turelli2017commentary}, \revpoint{1}{8}
this assumption is rather fragile --
a Gaussian distribution for the population only emerges under weak or Gaussian selection.
However, the model has been used in influential work.
In principle, another factor that could produce a non-Gaussian population distribution
is if the Mendelian sampling terms are not Gaussian, as we consider here --
so, it is interesting to look for other situations where the population distribution is analytically tractable.

@@

Since $\varphi$ and $\Phi$ are characteristic functions, they are continuous and $\phi(0) = \Phi(0) = 1$, and thus both are non-zero in a neighbourhood of 0. In particular, $\psi(\theta) = \Log{\varphi}(\theta)$ and $\Psi(\theta) = \Log{\Phi(\theta)}$ exist in some neighbourhood of 0 as well\footnote{We use $\Log$ to denote the principal value of the complex logarithm, for which the imaginary part always lies in the interval $(-\pi,\pi]$.}.  From the above, we have
\[
	\Log{F^{(n)}(\varphi)(\theta)} = {\textstyle 2^{n} \psi\left(\frac{\theta}{2^{n}}\right)} 
		+ \sum_{k=0}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}.
\]
We briefly explore what possible limiting values are valid characteristic functions. 
We refer to the first term on the left hand side as the \emph{reproduction term} and the remaining sum as the \emph{noise term}, and only consider the case \revpoint{2}{8}
when both have well-posed limits independent of the other. \llabel{limit_remark} % We don't need to limit to this caee.
Biologically,
the ``noise term'' is the contribution to variation in the population
generated by the Mendelian sampling terms (\ie, by recombination and mutation),
while the ``reproduction term'' is the contribution from variation in the initial generation.
\llabel{clarification2}

@@

\subsection*{Acknowledgements}
Thanks go to Nate Pope for useful discussion
and to Gregor Gorjanc for insight into modern uses
of the animal model and genomic selection. 
This project was started on a visit by TLP to the University of Oregon that was partially supported by the CNRS PEPS grant ``Jeunes Chercheuses et Jeunes Chercheurs''.

>>>>

\begin{abstract}
    The infinitesimal model of quantitative genetics
    relies on the Central Limit Theorem to stipulate that
    under additive models of quantitative traits
    determined by many loci having similar effect size,
    the difference between an offspring's genetic trait component
    and the average of their two parents' genetic trait components
    is Normally distributed and independent of the parents' values.
    Here, we investigate how the assumption of similar effect sizes affects the model:
    if, alternatively, the tail of the effect size distribution is polynomial
    with exponent $\alpha < 2$, then a different Central Limit Theorem implies that
    sums of effects should be well-approximated by a ``stable distribution'',
    for which single large effects are often still important.
    Empirically, we first find tail exponents between 1 and 2 in 
    effect sizes estimated by genome-wide association studies
    of many human disease-related traits.
    We then show that the independence of offspring trait deviations from parental averages
    in many cases implies the Gaussian aspect of the infinitesimal model,
    suggesting that non-Gaussian models of trait evolution must explicitly
    track the underlying genetics, at least for loci of large effect.
    We also characterize possible limiting trait distributions
    of the infinitesimal model with infinitely divisible noise distributions,
    and compare our results to simulations.
\end{abstract}

@@

The Gaussian distribution appears thanks to ``the'' Central Limit Theorem,
i.e., the universality of adding up a great many very small and independent things.
However, there is a wider class of Central Limit Theorems,
which describe the distributions of sums of a great many independent and exchangeable things:
the \emph{stable laws} \citep{gnedenko1968limit}.
Roughly speaking, these say that if we add together $n$ independent copies of a random variable $X$
for which $\P\{ X > t \}$ is proportional to $t^{-\alpha}$ for some $0 < \alpha < 2$,
then -- almost regardless of the details -- the result, scaled by $n^{1/\alpha}$,
is well-approximated by a universal form, an $\alpha$-stable distribution
(and the approximation becomes exact, as $n \to \infty$).
Oth the other hand, if $\alpha \ge 2$ then the distribution has finite variance, \revpoint{2}{1} 
and so the sum of $n$, divided by $\sqrt{n}$, converges to a Gaussian distribution.
In such ensembles, single large entries can be important:
concretely, the largest of the $n$ copies will be of order $n^{1/\alpha}$,
so that e.g., for $\alpha=1$ a positive proportion of the sum of $n$ entries will be contributed by a single one.
On the other hand, the largest of $n$ Gaussians is of order $\sqrt{\log n}$. \revpoint{1}{11}

This suggests exploring whether stable distributions might reasonably stand in for the Gaussian
in the infinitesimal model. \revpoint{1}{1}
In this paper, we make preliminary explorations in that direction:
Do the results of \citet{barton2017infinitesimal}
carry over when the distribution of effect sizes falls in the domain of attraction of a stable law?
The immediate answer to this will be ``no'' --
in fact, we will see that independence of offsprings' deviations and midparent values \revpoint{1}{2}
\emph{implies} a Gaussian distribution for the offsprings' deviations, in some sense.
This independence is a key reason why the infinitesimal model is tractable,
and so a conclusion is that additive models with non-Gaussian effects may not be well-approximated
by ``trait-only'' models that do not explicitly represent underlying genetic variation.
Having thus shown that ``infinitesimal models with non-Gaussian noise'' is on weak footing as a model of reality,
we then go on to analyze exactly these models.
(As motivation,
we have not shown that such models are \emph{not} good approximations to reality,
and it is interesting and perhaps informative to understand how the noise distribution affects results.)

@@

The ways in which genetic variants affect an organisms' trait can be complex,
and interactions between alleles, between loci, and with the environment are often important in practice.
In fact, some important traits that have been claimed to be mostly genetically determined \revpoint{1}{14}
only appear heritable through non-genetic means,
e.g., systematic inequalities in resource distribution partitioned by class or race
\citep[e.g.,][]{davenport1916hereditary}.
In this paper, we follow the literature in working with the relatively simple purely additive case
of a trait that can be decomposed into a sum across loci of the contributions of the alleles from each locus,
even omitting an independent effect of ``the environment''.
However, we emphasize that such models should not always be the default for understanding biological variation,
especially for human traits.

@@

Since the Mendelian sampling term is a sum of terms of the form
$$
    (X_\ell - 1/2) \eta_\ell^{[1]} + (1/2 - X_\ell) \eta_\ell^{[2]},
$$
where $\eta_\ell^{[i]}$ is the effect of the allele in parent $i$ at locus $\ell$,
the appropriate central limit theorem for the sum is determined by
the tail behavior of $|(X - 1/2) \eta|$,
where $X$ is Bernoulli(1/2) and $\eta$ is the effect
of a uniformly chosen one of the two alleles from a uniformly chosen parent
at a uniformly chosen locus.
Since $|X-1/2| = 1/2$, we'd like to know
if $\P\{|X|>2t\}$ is proportional to $t^{-\alpha}$,
i.e., the value of
$$
    \lim_{t \to \infty} \frac{ - \log \P\{ |X| > 2 t \} }{ \log t } = \alpha ,
$$
if the limit exists.
Let $p_\ell$ be the minor allele frequency at the $\ell^\text{th}$ locus;
since the data we have assigns effect 0 to the major allele,
$\eta_\ell = 0$ with probability $1-p_\ell$.
Therefore, if we have estimated effects for $M$ SNPs
with minor allele frequency $p_\ell$ and effect size $e_\ell$ for the minor allele,
for $1 \le \ell \le M$,
then
$$
    \P\{ |\eta| > 2t \} \approx \frac{1}{M} \sum_{\ell : |e_\ell| > 2t} p_\ell .
$$
Note that the relevant effect size distribution here is weighted by frequency, \revpoint{1}{3}
which is necessary because we're looking at the deviation of a typical offspring from their parents,
and the chance a given allele is carried by a random parent is given by the frequency.
(This weighting will make a difference, as frequencies and effect sizes are often correlated in practice.)
So, to obtain a per-trait estimate of $\alpha$,
we plot $\log \frac{1}{M} \sum_{\ell : |e_\ell| > t} p_\ell$ against $\log t$
for a sequence of values of $t$,
and fit a linear model to the tail.
(To do this,
we take as the ``tail''
all values of $t$ for which the number of SNPs with absolute effect above $t$
is at least 20 and at most 10\% of all SNPs.) \revpoint{2}{3}
Examples are shown in Figure~\ref{fig:example_snps}.

@@

A possible concern is that tail exponents are difficult to estimate accurately, \revpoint{1}{4}
but simulation-based estimates of accuracy (Supplemental Figure~\ref{fig:power})
show that our estimates are unbiased but only accurate to within roughly $\pm 0.4$.
So, the distribution of Figure~\ref{fig:exponent_hist} should not be taken as
a direct estimate of the underlying distribution of $\alpha$ values,
but is robust evidence for exponents strictly between 1 and 2 (and probably around 1.5).

@@

The Kac-Bernstein theorem states that if two independent random variables
have the property that their sum and difference are independent of each other,
then the random variables have a joint Gaussian distribution
(\citet{kac1939characterization,bernstein1941property},
and see \citep{kagan1973characterization} for many much more general results).
Since $R_M$ is (half) the difference and $\bar Z_M$ is (half) the sum
of $Z_M$ and $V_M$, and $Z_M$ and $V_M$ are independent,
if we assume that also $R_M$ and $\bar Z_M$ are independent,
then both are therefore jointly Gaussian.

@@

We should now discuss what this counter-intuitive result says and does not say.
The proposition says, roughly, that if the parental alleles are unlinked and independently chosen,
and meiotic segregation is fair,
then independence of Mendelian sampling term from midparent value
\emph{implies} a joint Gaussian distribution for the two.
We have not assumed anything about the distribution of the trait in the population:
rather, we have showed that any model of this sort
that assumes independence of Mendelian sampling term and midparent value
must also, by logical necessity, have a Gaussian distribution for the two.
In other words, \revpoint{1}{7}
since models can be easily set up for which offspring minus midparent traits
are not Gaussian,
the implication of the proposition
is that for such models, independence of the Mendelian sampling term
is not likely a good assumption.
Note that the proof does not depend on independence of the $X_\ell$:
indeed, alleles may be inherited together,
although in practice this would quickly create dependencies between the effects
of alleles at such linked loci. \revpoint{2}{5}
Also note that it is \emph{not} usually true that the components --
for instance, $U_M^{[1]}$ and $U_M^{[2]}$ --
are independent of each other.
Finally, note that although they may not be independent,
the Mendelian sampling term $R_M$ and the midparent value $\bar Z_M$
\emph{are} always uncorrelated, since $Z_M$ and $V_M$ are exchangeable.

@@

% NOT HELPFUL
% Finally, a possible point of confusion is that \revpoint{1}{6}
% the infinitesimal model is often phrased as a statement about offspring trait distributions
% \emph{given} parental trait values,
% while our statement above regards parental trait values as random.
% However, the notions are equivalent:
% the statement that
% ``for each pair of parents the deviation of the offspring's trait from midparent
% has a distribution that does not depend on the parents' values''
% is essentially the definition of statistical independence
% of offspring's Mendelian sampling term and parental trait values.

@@

As observed in \cite{turelli2017commentary},
there are at least three interpretations of what is called ``the infinitesimal model''. 
The maximalist of these, what \cite{turelli2017commentary} calls the Gaussian-population approximation,
assumes parental trait-independent Gaussian Mendelian sampling terms,
and, moreover, that  population trait values are in deterministic mutation-selection balance, with a Gaussian empirical distribution
(\ie the \emph{fraction} of the population with trait value in $[z,z+dz)$
is approximately $\frac{1}{2\pi\sigma^{2}} e^{-(z-m)^{2}/(2\sigma^{2})} dz$
for suitable choices of the empirical mean and variance $m$ and $\sigma^{2}$).
As also discussed in \citet{turelli2017commentary}, \revpoint{1}{8}
this assumption is rather fragile --
a Gaussian distribution for the population only emerges under weak or Gaussian selection.
% However, the model has been used in influential work.
In principle, another factor that could produce a non-Gaussian population distribution
is if the Mendelian sampling terms are not Gaussian, as we consider here --
so, it is interesting to look for other situations where the population distribution is analytically tractable.

@@

Since $\varphi$ and $\Phi$ are characteristic functions, they are continuous and $\phi(0) = \Phi(0) = 1$, and thus both are non-zero in a neighbourhood of 0. In particular, $\psi(\theta) = \Log{\varphi}(\theta)$ and $\Psi(\theta) = \Log{\Phi(\theta)}$ exist in some neighbourhood of 0 as well\footnote{We use $\Log$ to denote the principal value of the complex logarithm, for which the imaginary part always lies in the interval $(-\pi,\pi]$.}.  From the above, we have
\[
	\Log{F^{(n)}(\varphi)(\theta)} = {\textstyle 2^{n} \psi\left(\frac{\theta}{2^{n}}\right)} 
		+ \sum_{k=0}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}.
\]
We briefly explore what possible limiting values are valid characteristic functions. 
We refer to the first term on the left hand side as the \emph{reproduction term} and the remaining sum as the \emph{noise term}, and only consider the case \revpoint{2}{8}
when both have well-posed limits independent of the other. \llabel{limit_remark} % We don't need to limit to this caee.
Biologically, \llabel{clarification2}
the ``noise term'' is the contribution to variation in the population
generated by the Mendelian sampling terms (\ie, by recombination and mutation),
while the ``reproduction term'' is the contribution from variation in the initial generation.

@@

\subsection*{Acknowledgements}
Thanks go to Nate Pope for useful discussion
and to Gregor Gorjanc for insight into modern uses
of the animal model and genomic selection,
as well as to two anonymous reviewers for helpful comments.
This project was started on a visit by TLP to the University of Oregon that was partially supported by the CNRS PEPS grant ``Jeunes Chercheuses et Jeunes Chercheurs''.
